---
title: "R Notebook"
output: html_notebook
---

# Notes in R by Eric

## Declare variables

Let's declare variables 'a' and 'b' and calculate their sum. We will obtain the result in the variable 'suma'.

```{r}
a<-5
b<-10
suma<-a+b
suma
```

## Type of data

Let's display the type of data from our columns

```{r}
str(mtcars)
```

Let's display the data type of a column and convert it from integer to boolean.

```{r}
class(mtcars$vs)
class(mtcars$am)
mtcars$vs = as.logical(mtcars$vs)
mtcars$am = as.logical(mtcars$am)
str(mtcars)
```

## Measures of central tendency

Let's display the measures of central tendency

```{r}
summary(mtcars$wt)#just display data a column
summary(mtcars)#to display all columns
```

## Calculate a column

Let's convert from pounds to kilograms in column wt in new datasest

```{r}
mtcars.ds1 = transform(mtcars,wt=(wt*1000)/2)
mtcars.ds1
```

## The vectors

this declare with c()

```{r}
horas_de_suenio <- c(8,7,7,7,7)
adicionales_suenio <- c(1,2,2,2,2)
ficcion_de_suenio<- horas_de_suenio + adicionales_suenio
ficcion_de_suenio
```

## Matrix

let's will construct the matrix from the vectors how example. just sum our. and Now let´s add a row to our matrix.

```{r}
nombre_titulos<-c("Lunes","Martes","Miercoles","Jueves","Viernes")
nombre_filas<-c("horas de sueño","adicionales de sueño")
horas_de_suenio <- c(8,7,7,7,7)
adicionales_suenio <- c(1,2,2,2,2)
horas_totales<-matrix(c(horas_de_suenio,adicionales_suenio),nrow = 2,byrow = TRUE)
colnames(horas_totales)<-nombre_titulos
rownames(horas_totales)<-nombre_filas

horas_totales
colSums(horas_totales)

final_matriz<-rbind(horas_totales,c(1,1,1,1,2))
final_matriz

colSums(horas_totales)
```

## Comparison operator

Which are

```{r}
1==1#equial to
1!=1#not equal to
1<1 #less than 
1<=1#less than or equal to 
1>1 #greater than
1>=1#greater than or equal to
1|1 # or
1&1 #and
valor <- !1 #No
valor
1%in%1 #in dataset
```

Example of how to use in dataset if you want to filter a columns

```{r}
mtcars[mtcars$cyl>=6 & mtcars$vs==FALSE,]
```

we create a new subset.

syntax: subset(dataset, condition, columns)

```{r}
mtcars_ss<-subset(mtcars,mtcars$cyl>=6 & mtcars$vs==FALSE,select = c(model,cyl))
mtcars_ss
```

rename column

```{r}
colnames(mtcars_ss)[colnames(mtcars_ss) == "model"] <- "modelo"
mtcars_ss
```

## Factor

```{r}
Nivel <- c("basico","intermedio","avanzado")
Nivel
```

```{r}

datos <- c("manzana", "naranja", "manzana", "plátano", "naranja")
factor_datos <- factor(datos)
levels(factor_datos)
```

## List

```{r}
el_vector<- 1:8
la_matiz<-matrix(1:9, ncol=3)
el_df<- mtcars[1:3]
la_lista<-list(el_vector,la_matiz,el_df)
la_lista
```

## How to see dataset another way

just title and rows number

```{r}
head(mtcars, n=2)
```

rows end

```{r}
tail(mtcars,n=2)
```

also to see the dataset

```{r}
str(mtcars)
```

# EDA

Exploratory Data Analysis

## Scatter plox

```{r}
plot(mtcars$mpg,mtcars$hp,
      xlab="cilindros",ylab = "caballos de fuerza",
     main = "Relacion cilindros y milla por galon")
```

## Histogram

```{r}
hist(mtcars$hp, breaks="Sturges", main= "Histogram sturges", xlab="valores", ylab="frecuencia")
hist(mtcars$hp, breaks="Scott", main= "Histogram scott", xlab="valores", ylab="frecuencia")
```

with ggplot2

```{r}
library(ggplot2)
# histograsma mtcars qplot
library(ggplot2)

ggplot(mtcars, aes(x = hp)) + 
  geom_histogram(binwidth = 10) + xlab("caballos de fuerza") + 
  ylab("carros") + 
  ggtitle("Histogram") +
  theme(legend.position = "none") +
  theme(panel.background = element_blank(),panel.grid.major = element_blank())
```
